Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-04-10 paper/paper.tex
Update tex.

The Java ASTs in our analysis were generated using Spoon~\cite{pawlak:hal-01169705}. Spoon is a tool for transforming and analysing Java source code. It breaks code up into a meta-model including the AST itself. The model consists of three parts: structural elements, code elements and references to program elements. According to Spoon's authors: ''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.  The code part contains the executable Java code, such as the one found in method bodies. The reference part models the references to program elements (for instance a reference to a type)''~\cite{pawlak:hal-01169705}. The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards. Spoon also preserves all type and package (library) information—information vital to our semantic analysis.

@@

Once we have the generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.

@@

@@

This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted. Note the fully qualified type names: the source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}. Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.

In order to gather all the changes for a given repository over time we generated ASTs for each commit and compared the difference. We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}). Each commit and its parent were then compared, using AST differencing as described above. There are a few edges cases, however: If a commit has no parent, then every declaration or invocation in the commit was treated as an addition. Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}). If a commit has more than one parent, this indicates a merge commit in Git. Merge commits contain the changes made in each parent commit combined. Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.

@@

The analyzed data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}). The source code is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.

TypeV provide a variety of views to help organize the data and provide a finer level of detail. The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}). The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub's code frequency graph (Figure~\ref{fig:GH_code_freq}). The height of each bar shows the number of changes to a type in a smaller unit of time. This is calculated by multiplying the bar height by the number of changes during that time period divided by the maximum number of changes attested for this type. A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods. The other bars in the row are proportional to that maximum number. The width of each bar represents a set time period and can be changed to an hour, day, week, or month. The rows are sorted so that the graph for the most used types appear the top and are in descending order.

>>>>

The Java ASTs in our analysis were generated using Spoon~\cite{pawlak:hal-01169705}. Spoon is a tool for transforming and analyzing Java source code. It breaks code up into a meta-model including the AST itself. The model consists of three parts: structural elements, code elements and references to program elements. According to Spoon's authors: ''The structural part contains the declarations of the program elements, such as interface, class, variable, method, annotation, and enum declarations.  The code part contains the executable Java code, such as the one found in method bodies. The reference part models the references to program elements (for instance a reference to a type)''~\cite{pawlak:hal-01169705}. The Spoon model is convenient because it provides the infrastructure for performing AST differencing (described later) while retaining the code elements (such as type information) for analysis afterwards. Spoon also preserves all type and package (library) information, information vital to our semantic analysis.

@@

Once we generated ASTs, we were able to traverse through the trees to count the number of added or deleted \emph{declarations} and \emph{invocations} for a given type.

@@

This states that two declarations of type \texttt{java.net.URI} were added, and two declarations of type \texttt{java.lang.String} were deleted. Note the fully qualified type names. The source text may not literally have the characters \texttt{java.net.URI}, but Spoon was able to infer from the import statements in the Java source text that references to ''URI'' resolve to the fully-qualified type name \texttt{java.net.URI}. Thus, TypeV is able to keep track of the actual class being used and is not confused when two classes in separate packages have the same name.

@@

In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference. We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}). Each commit and its parent were then compared using AST differencing as described above. There are a few edge cases; however, if a commit has no parent, then every declaration or invocation in the commit was treated as an addition. Since we compared each commit to its parent, as reported by Git, we were able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}). If a commit has more than one parent, this indicates a merge commit in Git. Merge commits contain the changes made in each parent commit combined. Therefore, we ignored merge commits since all changes in the merge should already have been accounted for in previous commits.

@@

The data would be difficult to analyze if displayed as its raw text form; instead, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}). The source code for TypeV is available online on GitHub,\footnote{\url{https://github.com/mdfeist/TypeV}}.

TypeV provide a variety of views to help organize the data and provide different levels of detail. The main view displays a series of stacked bar graphs which show changes to types over a period of time in a project (Figure~\ref{fig:main-view}). The blue and orange portions account for the number of additions and deletions in a similar fashion to GitHub’s code frequency graph (Figure~\ref{fig:GH_code_freq}). The height of each bar shows the number of changes to a type. The width of each bar represents a set time period and can be changed to an hour, day, week, or month. A bar which fills the row vertically is the period with the largest number of changes to that type out of the visible time periods. The other bars in the row are proportional to that maximum number.  The bar height is calculated by multiplying the maximum bar height by the number of changes during that time period divided by the maximum number of changes attested for this type. The rows are sorted so that the graph for the most used types appear the top and are in descending order.
